---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-figs/"
)
```

# Stock-assessment methods for data-limited fisheries

[![Build Status](https://magnum.travis-ci.com/datalimited/datalimited.svg?token=QExyQi6ySw3SZD4gggYN&branch=master)](https://magnum.travis-ci.com/datalimited/datalimited)

## Installation

Before installing datalimited, you will need to [install JAGS](http://mcmc-jags.sourceforge.net) and have a [C++ compiler set up](https://support.rstudio.com/hc/en-us/articles/200486498-Package-Development-Prerequisites).

The R package datalimited can then be installed from GitHub with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("datalimited/datalimited")
```

The package implements the methods used in Rosenberg et al. (2014) including the following:

- Catch-MSY based on Martell and Froese (2013); see `?cmsy`
- Catch-only method with sample importance resampling based on asconcellos and Cochrane (2005); see `?comsir`
- Panel regression in the style of Costello et al. (2012); see `?prm`
- State-space catch-only model from Thorson et al. (2013); see `?sscom`

## An example with `cmsy()`

```{r cmsy}
library("datalimited")
set.seed(1)
x <- cmsy(blue_gren$yr, ct = blue_gren$ct, reps = 2e4)
names(x)
par(mfrow = c(2, 2))
plot(blue_gren$yr, blue_gren$ct, type = "o", xlab = "Year", 
  ylab = "Catch (t)")
plot(blue_gren$yr,  apply(x$biomass, 2, median)[-1], type = "o",
  ylab = "Estimated biomass", xlab = "Year")
hist(x$bmsy)
plot(x$theta$r, x$theta$k, col = "#00000030")
library("ggplot2")
ggplot(x$bbmsy, aes(year, bbmsy_q50)) + geom_line()  +
    geom_ribbon(aes(ymin = bbmsy_q25, ymax = bbmsy_q75), alpha = 0.2) +
    geom_ribbon(aes(ymin = bbmsy_q2.5, ymax = bbmsy_q97.5), alpha = 0.1) +
    geom_hline(yintercept = 1, lty = 2) + theme_light()
```

## Citation

To cite the package, cite the authors of the original assessment method (shown below), Rosenberg et al. (2014), and:

```{r}
citation("datalimited")
```


## References

Costello, C., D. Ovando, R. Hilborn, S. D. Gaines, O. Deschenes, and S. E. Lester. 2012. Status and solutions for the world’s unassessed fisheries. Science 338:517-520.

Martell, S., and R. Froese. 2013. A simple method for estimating MSY from catch and resilience. Fish and Fisheries 14:504-514.

Rosenberg, A. A., M. J. Fogarty, A. B. Cooper, M. Dickey-Collas, E. A. Fulton, N. L. Gutiérrez, K. J. W. Hyde, K. M. Kleisner, C. Longo, C. V. Minte-Vera, C. Minto, I. Mosqueira, G. C. Osio, D. Ovando, E. R. Selig, J. T. Thorson, and Y. Ye. 2014. Developing new approaches to global stock status assessment and fishery production potential of the seas. FAO Fisheries and Aquaculture Circular, Rome, Italy.

Thorson, J. T., C. Minto, C. V. Minte-Vera, K. M. Kleisner, and C. Longo. 2013. A new role for effort dynamics in the theory of harvested populations and data-poor stock assessment. Canadian Journal of Fisheries and Aquatic Sciences 70:1829–1844.

Vasconcellos, M., and K. Cochrane. 2005. Overview of World Status of Data-Limited Fisheries: Inferences from Landings Statistics. Pages 1-20 in G. H. Kruse, V. F. Gallucci, D. E. Hay, R. I. Perry, R. M. Peterman, T. C. Shirley, P. D. Spencer, B. Wilson, and D. Woodby, editors. Fisheries Assessment and Management in Data-Limited Situations. Alaska Sea Grant, University of Alaska Fairbanks.
